---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r global, echo = FALSE, results = "hide", message = FALSE }
library(data.table)
library(ggplot2)
library(lubridate)
```


## Loading and preprocessing the data

```{r load_data, results = "hide" }
unzip("activity.zip")
activity <- fread("activity.csv", sep = ",", na.strings = "NA")
activity[, date := as.POSIXct(date)]
```

## What is mean total number of steps taken per day?

```{r mean_total_number_histogram, results = "hide" }
daily_steps <- activity[, list(steps = sum(steps, na.rm = TRUE)), by = date]

ggplot(daily_steps, aes(x = steps)) + 
  geom_histogram(binwidth = 500, colour="black", fill="white") +
  geom_vline(aes(xintercept=mean(steps, na.rm=T)),
             color="red", linetype="dashed", size=1) +
  geom_vline(aes(xintercept=median(steps, na.rm=T)),
             color="blue", linetype="dashed", size=1)
```

```{r mean_total_number}
mean_total_number <- mean(daily_steps$steps)
median_total_number <- median(daily_steps$steps)
```

The mean total number of steps taken per day is `r mean_total_number` and the median is `r median_total_number`.

## What is the average daily activity pattern?

```{r average_daily_activity_patern }
interval_steps_mean <- activity[, list(steps_mean = mean(steps, na.rm = TRUE)), by = interval]
plt <- ggplot( data = interval_steps_mean, aes( interval, steps_mean )) + 
  geom_line()
print(plt)

interval_with_max_steps <- interval_steps_mean[,.SD[which.max(steps_mean)]]
```

The interval with max number of step is `r interval_with_max_steps$interval`.

## Imputing missing values

```{r number_missing_values }
missing_steps <- sum(is.na(activity$steps))
```

There are `r missing_steps` missing values (i.e. intervals without steps).

```{r fill_missing_values, results = "hide" }
activity_filled <- merge(activity, interval_steps_mean, by = "interval")
activity_filled[is.na(steps), steps := as.integer(round(steps_mean))]
daily_steps_filled <- activity_filled[, list(steps = sum(steps, na.rm = TRUE)), by = date]

ggplot(daily_steps_filled, aes(x = steps)) + 
  geom_histogram(binwidth = 500, colour="black", fill="white") +
  geom_vline(aes(xintercept=mean(steps, na.rm=T)),
             color="red", linetype="dashed", size=1) +
  geom_vline(aes(xintercept=median(steps, na.rm=T)),
             color="blue", linetype="dashed", size=1)
```

```{r filled_mean_total_number, results = "hide"}
mean_total_number_filled <- mean(daily_steps_filled$steps)
median_total_number_filled <- median(daily_steps_filled$steps)
```

With empty values replaced by the mean value of empty interval, the total mean and total median values are changed. Before the empty values are replaced, the mean was `r mean_total_number` and median was `r median_total_number`. Once empty variables replaced the mean value is `r mean_total_number_filled` and the median is `r median_total_number_filled`.


## Are there differences in activity patterns between weekdays and weekends?

```{r weekday_pattern, results = "hide" }
activity[wday(date) %in% c(1,7), weekday := "weekend"]
activity[wday(date) %in% c(2,3,4,5,6), weekday := "weekday"]
interval_steps_mean <- activity[, list(steps_mean = mean(steps, na.rm = TRUE)), 
                                by = .(weekday, interval)]

ggplot( data = interval_steps_mean, aes( interval, steps_mean )) +  
      facet_grid(.  ~	weekday) + 
      geom_line()
```
